---
title: "U.S. Forest Fires"
author: "Juan Carlos (juancg@uw.edu), Chantalle Matro (cmatro@uw.edu), Militha Madur (militham@uw.edu)"
date: "11/14/2022"
output: html_document
---

library(knitr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(plotly)

```{r setup, include=FALSE}
source("../source/summary_info.R")
source("../source/table_info.R")
source("../docs/Chart1.R")
source("../docs/Chart2.R")
source("../docs/Chart3.R")
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Overview

Below is a report that summarizes forest fires in the United States from 1900-2021. Two different data sets were analyzed - U.S. Fire Origins ([source](https://redivis.com/datasets/5k9t-07xsg7ckc/tables/5r3s-4s2avr2eb?variable=initial_response)) and U.S. Fire Perimeters ([source](https://redivis.com/datasets/5k9t-07xsg7ckc/tables/9r58-8ckedbnhw)). Over this time period, key insights included the following. The most common statistical cause of forest fires in the U.S. was lightning. The greatest number of fires occurred in 2017. Most fires originated in California. The maximum total acreage a forest fire spanned was 3,611,520 acres, while the most common fire size classification was level A, which corresponds to .00-0.25 acres. Additionally, this report includes three different visualizations that break down each of these parameters specifically.

## Table

### Table 1 - U.S. Fire Origin Points Summary

```{r echo = FALSE}
table_info
```
### Table 2 - U.S. Fire Perimeters Summary

```{r echo = FALSE}
table_info_2
```

These tables summarize the most relevant data from the U.S. Fire Origin Points and U.S. Fire Perimeters data sets. For each state available on the former data set, their respective counties which have experienced the most fires were retrieved. The table also shows the most common fire size and causes for each state. The latter data set was grouped by the year fires occurred, and the largest total acres and common causes were logged.

## Charts

### Chart 1 - Mapping the locations/origin points of Forest Fires in the U.S.

Here is the map:

```{r echo = FALSE}
fig
```

Chart one is a map of the origin points and size classifications of forest fires in the United States from 1990-2021. This date range was chosen in order to focus future analysis on the effects of climate change on the origin and spread of forest fires. The location of each point represents the origin point of the forest fire and the corresponding color identifies the forest fire size classification of the fire. The purpose of this graph is to map the most frequent locations/regions of forest fires as well as the range of fire sizes. The region with the greatest number of forest fire origin points is the Western United States, more specifically the West Coast (California). Additionally, the Western United States has the greatest number of forest fires that are classified as G in forest fire size, which is greater than 5000 acres. The map also shows that the origin points of forest fires that affect the United States are not limited to starting in the United States. Many forest fires originated in Canada, Mexico, and nearby islands.

### Chart 2 - Visualization for the most frequent cause on BarChart

Here is the bar graph:

```{r echo = FALSE}
graph
```

Chart two is a bar graph that shows the frequency of the causes of forest fires in the United States. The purpose of this chart is to see what the biggest threats are to starting the forest fires. The biggest cause that was shown was lightning. Lightning was much greater than even the second biggest cause, which was camp fires. The two lowest causes were railroad and children.

### Chart 3 - Visualization for the density of fires for every year

Here is the graph:

```{r echo = FALSE}
plot
```

Chart three is a density chart of the fires in every year. This chart was used to show the trend of the density of fires per year since 1900 until up to 2021. From this density chart you can see that the density of fires per year has been increasing year by year. In 1925 the density of fires was around .004 and now it is at a little above .020. The density of fires per year since 1925 has increased by more than four times.
